---
title: "Giardia rectal swab analysis"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

Setup.
```{r setup, message=FALSE, warning=FALSE}
#Install packages
#install.packages("xtable", repos="http://cran.us.r-project.org")
#install.packages("scales", repos="http://cran.us.r-project.org")
#install.packages("bazar", repos="http://cran.us.r-project.org")
#install.packages("gmodels", repos="http://cran.us.r-project.org")
#install.packages("stargazer", repos="http://cran.us.r-project.org")
#install.packages("cowplot", repos="http://cran.us.r-project.org")

#Load packages
library(tidyverse)
library(lubridate)
library(epiR)
library(xtable)
library(scales)
library(bazar) #sumNA function
library(gmodels)
library(stargazer)
library(cowplot)
library(ggthemes)

#Color palette
library(RColorBrewer)
#display.brewer.pal(9, name="BrBG")
BrBG6 <- c("#8C510A" ,"#BF812D", "#DFC27D", "#80CDC1", "#35978F", "#01665E")
BrBG5 <- c("#8C510A" ,"#BF812D", "#DFC27D", "#35978F", "#01665E")
BrBG4_brown <- c("#f6e8c3", "#dfc27d", "#bf812d", "#8c510a")
BrBG4_turq <- c("#c7eae5", "#80cdc1", "#35978f", "#01665e")
BrBG4_turq_dark <- c("#80cdc1", "#35978f", "#01665e", "#003c30")
BrBG2 <- c("#003c30", "#35978f")


#Working directorories (set one to comment depending on who is working on the file)
#Kristen: 
#setwd("~/Dropbox/_Papers/Rectal swab STH/Analysis/")

#Jacqueline
setwd("/Users/razel/Dropbox/Rectal swab STH/Analysis/")

```

Load data. Merging was done in Stata.
```{r data}

d <- (read.csv("DataPrep/GiardiaPCRStoolRectalMerged_03.04.19.csv") %>%
        mutate(dob_mcd = as.Date(dob_mcd, "%m/%d/%Y"),
               visitdate = as.Date(appcreateddatess, "%m/%d/%Y"),
               ageM = as.numeric(floor((visitdate - dob_mcd)/30)),
               ageY = as.numeric(floor((visitdate - dob_mcd)/365)),
               bsfs1 = fct_recode(sampleconsistencygrader1, "Loose" = "Fluffy With Ragged Edges", 
                                  "Smooth" = "Smooth Soft", "Pellet" = "Separate Hard Lumps"),
               bsfs1 = fct_relevel(bsfs1, "Watery", "Loose", "Smooth", "Lumpy", "Pellet"),
               loosewat1 = fct_collapse(bsfs1,
                                       "loose/watery" = c("Watery", "Loose"),
                                       "formed" = c("Lumpy", "Pellet", "Smooth"))) %>%
        select(- visiblestoolonswab1, -visiblestoolonswab2, -moisturein15mltube1, -moisturein15mltube2) %>%
        rename(gender = genderes, visiblestoolonswab1=visiblestoolonswab3, visiblestoolonswab2 = visiblestoolonswab4, 
               moisturein15mltube1=moisturein15mltube3, moisturein15mltube2=moisturein15mltube4, ctstool1 = ctmean1, 
               ctstool2 = ctmean2, ctrectal1 = ctmean3, ctrectal2 = ctmean4, vilID = villageeswuha)
               ) %>%
        mutate(condensation = ifelse(moisturein15mltube1=="Yes", 1, 0)) %>%
        rowwise() %>%
          mutate(eitherStoolPos = sumNA(stool1_pos, stool2_pos, na.rm = T),
                 eitherStoolPos = ifelse(eitherStoolPos >1, 1, eitherStoolPos),
                 eitherRectalPos = sumNA(rectal1_pos, rectal2_pos, na.rm = T), 
                 eitherRectalPos = ifelse(eitherRectalPos >1, 1, eitherRectalPos),
                 truePos = sumNA(stool1_pos, stool2_pos, rectal1_pos, rectal2_pos, na.rm = T), 
                 truePos = ifelse(truePos > 1, 1, truePos)
                 )

#Remove data points with low IAC values
d <- d %>%
  filter(stool1_ID != 491170 & stool1_ID != 460864)

#Mean CT
d <- d %>%
  mutate(ctstoolMax = max(ctstool1, ctstool2, na.rm=T),
         ctstoolMax = ifelse(is.na(ctstool1) & is.na(ctstool2), NA, ctstoolMax),
         ctstoolMean = mean(c(ctstool1, ctstool2)),
         #if only 1 ct value, replace mean with max of that value
         ctstoolMean = ifelse(is.na(ctstoolMean), ctstoolMax, ctstoolMean)
         )
```

Intra-test reliability (kappa statistics).
```{r reliability}
#Reliability Stool vs. stool
############################
#Step 1: create a confusion matrix
tab1 <- xtabs(~stool2_pos + stool1_pos, data=d)
tab1

#Step 2: reverse column order
tab1 <- tab1[ 2:1, 2:1]

#Step 3: use the epi test command to calculate sensitivity and specificity
epi.kappa(tab1)


#Reliability Swab vs. swab
############################
#Step 1: create a confusion matrix
tab2 <- xtabs(~rectal2_pos + rectal1_pos, data=d)
tab2

#Step 2: reverse column order
tab2 <- tab2[ 2:1, 2:1]

#Step 3: use the epi test command to calculate sensitivity and specificity
epi.kappa(tab2)

#All positives
tab3 <- xtabs(~eitherRectalPos + eitherStoolPos, data=d)
tab3 <- tab3[ 2:1, 2:1]
tab3 
epi.kappa(tab3)
```

Positive / negative agreement of qPCR outcomes for DNA isolated from rectal swabs and bulk stool. 
```{r agreement, results = "asis"}
#################################
# (+/-) Agreement tables
#################################

#ALL SWABS VS ALL STOOL
AgreementTab  <- addmargins(table(d$eitherRectalPos, d$eitherStoolPos, dnn=c("Rectal Swab","Bulk Stool Sample")))
colnames(AgreementTab) <- c( "Stool --", "Stool +", "Total")
rownames(AgreementTab) <- c( "Swab --", "Swab +", "Total")
AgreementTab 

print(xtable(AgreementTab), include.rownames=TRUE, booktabs = TRUE)

#In stargazer
agree <- as.data.frame(xtable(AgreementTab))
stargazer(agree, type = "latex", summary = FALSE, rownames = TRUE, 
          title = "PCR detection agreement by specimen type")

#In xtable
agree <- as.data.frame(xtable(AgreementTab))
print(xtable(agree, caption = "PCR detection agreement by specimen type"), 
      caption.placement = getOption("xtable.caption.placement", "top"))

#STOOL REPLICATE 1 VS STOOL REPLICATE 2
AgreementStool  <- addmargins(table(d$stool1_pos, d$stool2_pos, dnn=c("Bulk Stool Sample 1","Bulk Stool Sample 2")))
colnames(AgreementStool) <- c( "Stool 1  --", "Stool 1 +", "Total")
rownames(AgreementStool) <- c( "Stool 2  --", "Stool 2 +", "Total")

AgreementStool

print(xtable(AgreementStool), include.rownames=TRUE, booktabs = TRUE)

AgreementStool <- as.data.frame(xtable(AgreementStool))

stargazer(AgreementStool, type = "latex", summary = FALSE, 
          rownames = TRUE, title = "PCR detection agreement by bulk stool replicate group", 
          notes = "κ = 0.91 (95% CI 0.63–1.00)")

#SWAB REPLICATE 1 VS SWAB REPLICATE 2
AgreementSwab  <- addmargins(table(d$rectal1_pos, d$rectal2_pos, dnn=c("Rectal swab 1","Rectal swab 2")))
colnames(AgreementSwab) <- c( "Swab 1  --", "Swab 1 +", "Total")
rownames(AgreementSwab) <- c( "Swab 2  --", "Swab 2 +", "Total")

print(xtable(AgreementSwab), include.rownames=TRUE, booktabs = TRUE)

AgreementSwab <- as.data.frame(xtable(AgreementSwab))

stargazer(AgreementSwab, type = "latex", summary = FALSE, 
          rownames = TRUE, title = "PCR detection agreement by rectal swab replicate group", 
          notes = "κ = 0.79 (95% CI 0.51–1.00)")
```

Sensitivity, specificity, PPV, and NPV.
```{r epiTests}
######################################
#Calculate sensitivity and specificity
######################################

#Step 1: create a cross tabulation (also called a confusion matrix)
swabTabAll <- xtabs(~eitherRectalPos + truePos, data=d)
swabTab1 <- xtabs(~rectal1_pos + truePos, data=d)
swabTab2 <- xtabs(~rectal2_pos + truePos, data=d)
stoolTabAll <- xtabs(~eitherStoolPos + truePos, data=d)
stoolTab1 <- xtabs(~stool1_pos + truePos, data=d)
stoolTab2 <- xtabs(~stool2_pos + truePos, data=d)

#Step 2: reverse column order
swabTabAll <- swabTabAll[2:1, 2:1]
swabTabAll

swabTab1 <- swabTab1[2:1, 2:1]
swabTab1

swabTab2 <- swabTab2[2:1, 2:1]
swabTab2

stoolTabAll <- stoolTabAll[2:1, 2:1]
stoolTabAll

stoolTab1 <- stoolTab1[2:1, 2:1]
stoolTab1

stoolTab2 <- stoolTab2[2:1, 2:1]
stoolTab2

#Step 3: use the epi test command to calculate swab sensitivity and specificity WITH NA's REPLACED WITH 0's
epi.tests(swabTabAll)
#sensitivity with all swab positives vs truePos = 68%
#specificity with all swab positives vs truePos = 100%
epi.tests(swabTab1)
#sensitivity with swab 1 positives vs truePos = 63%
#specificity with swab 1 positives vs truePos = 100%
epi.tests(swabTab2)
#sensitivity with swab 2 positives vs truePos = 69%
#specificity with swab 2 positives vs truePos = 100%
epi.tests(stoolTabAll)
#sensitivity with all stool positives vs truePos = 96%
#specificity with all stool positives vs truePos = 100%
epi.tests(stoolTab1)
#sensitivity with stool 1 positives vs truePos = 95%
#specificity with stool 1 positives vs truePos = 100%
epi.tests(stoolTab2)
#sensitivity with stool 2 positives vs truePos = 94%
#specificity with stool 2 positives vs truePos = 100%


#restrict to swabs with visable stool
####################################
table(d$visiblestoolonswab1)
#only 16 swabs with visable stool in rep 1 and 18 in rep 2

tabVisible <- xtabs(~rectal1_pos + truePos, data=subset(d, visiblestoolonswab1=="Yes"))
tabVisible
tabVisible <- tabVisible[2:1, 2:1]
epi.tests(tabVisible)
#sensitivity with swab 1 positives vs truePos = 60%
#specificity with swab 1 positives vs truePos = 100%


#restrict to watery/loose stool
#################################
tabLoose <- xtabs(~rectal1_pos + truePos, data=subset(d, loosewat1=="loose/watery"))
tabLoose
tabLoose <- tabLoose[2:1, 2:1]
epi.tests(tabLoose)
#sensitivity with all swab positives vs truePos = 50%
#specificity with all swab positives vs truePos = 100%


#restrict to condensation
#################################
tabCond <- xtabs(~rectal1_pos + truePos, data=subset(d, condensation==0))
tabCond
tabCond <- tabCond[2:1, 2:1]
epi.tests(tabCond)
#sensitivity with swab 1 positives vs truePos = 62%
#specificity with swab 1 positives vs truePos = 100%


#restrict you children (<12 months)
#################################
tab12 <- xtabs(~rectal1_pos + truePos, data=subset(d, ageM <12))
tab12
tab12 <- tab12[2:1, 2:1]
epi.tests(tab12)
#no effect here
#sensitivity with swab 1 positives vs truePos = 67%
#specificity with swab 1 positives vs truePos = 100%
```

Covariate effects.
```{r covariate, results = "asis"}
######################################################
#Model evaluating effects of age, 
#visable detection of stool and
#stool consistancy and condensation on sensitivity 
######################################################

d2 <- d %>%
  filter(truePos==1)

sens <- glm(eitherRectalPos~1 + ageM + visiblestoolonswab1 + loosewat1 + condensation, 
            family = binomial(link="logit"), data=d2)

coef(summary(sens))

sens2 <- glm(rectal1_pos~1 + ageM + visiblestoolonswab1 + loosewat1 + condensation, 
             family = binomial(link="logit"), data=d2)

coef(summary(sens2))


#Model evaluating effects of age, visable detection of stool, and stool consistancy
print(xtable(sens, 
             caption = "Effects of age, visable detection of stool, and stool consistancy on rectal swab sensitivity"),
             caption.placement = getOption("xtable.caption.placement", "top"))


#STARGAZER TABLE: specificity, sensitivity - soiling, loose/watery, condensation, age
sensSpec <- (read.csv("/Users/razel/Dropbox/Rectal swab STH/Analysis/DataPrep/JAM_CSV_4-29-19/swab_sensitivity_specificity_transposed.csv"))

colnames(sensSpec) <- c("", "Total (N = 57)", "Loose or watery (N = 11)", 
                        "Condensation (N = 75)", "Soiled (N = 15)", "< 12 months (N = 11)")

stargazer(sensSpec, summary = FALSE, rownames = FALSE, 
          title = "Covariate effects on sensitivity, specificity, and predictive value of rectal swabs for molecular detection of {\\emph{Giardia duodenalis}}")

#STARGAZER TABLE: specificity, sensitivity SIMPLIFIED AND TRANSPOSED
sensSpec <- (read.csv("/Users/razel/Dropbox/Rectal swab STH/Analysis/DataPrep/JAM_CSV_4-29-19/swab_sensitivity_specificity.csv"))

colnames(sensSpec) <- c("", "Swab sensitivity (95% CI)", "Swab specificity (95% CI)", "p-value")

stargazer(sensSpec, summary = FALSE, rownames = FALSE, 
          title = "Covariate effects on sensitivity and specificity of rectal swabs for the real-time PCR detection of Giardia duodenalis")
```

LaTex tables for paper. See LaTex document for final drafts.
```{r latex, results = "asis"}
#Stool summary
stoolCtSummary <- d %>%
  select(personID, ctstool1, ctstool2)

print(xtable(stoolCtSummary), include.rownames=TRUE, booktabs = TRUE)
stoolCtSummary <- as.data.frame(xtable(stoolCtSummary))
stargazer(stoolCtSummary, type = "latex", summary = TRUE, rownames = TRUE)

#Swab summary
swabCtSummary <- d %>%
  select(personID, ctrectal1, ctrectal2)

print(xtable(swabCtSummary), include.rownames=TRUE, booktabs = TRUE)
swabCtSummary <- as.data.frame(xtable(swabCtSummary))
stargazer(swabCtSummary, type = "latex", summary = TRUE, rownames = TRUE)

#POSITIVITY / MEAN CT TABLE
replicateSummary <- read_csv("/Users/razel/Dropbox/Rectal swab STH/Analysis/LaTex and CSV tables/giardia_table2.csv")

stargazer(replicateSummary, type = "latex", summary = FALSE, 
          rownames = FALSE, title = "PCR positivity and mean CT value by specimen replicate group")

# FALSE NEGATIVE / SENSITIVITY TABLE
sensitivitySummary <- read_csv("/Users/razel/Dropbox/Rectal swab STH/Analysis/LaTex and CSV tables/giardia_table5.csv")

stargazer(sensitivitySummary, type = "latex", summary = FALSE, 
          rownames = FALSE, title = "PCR negativity and diagnostic sensitivity by specimen replicate group")

# SEQUENCES TABLE
seqTable <- read_csv("/Users/razel/Dropbox/Rectal swab STH/Analysis/LaTex and CSV tables/giardia_sequences_table.csv")

stargazer(seqTable, type = "latex", summary = FALSE, rownames = FALSE, 
          title = "DNA target, primer, and probe sequences for the real-time PCR detection of Giardia duodenalis")

#write.csv(d, file = "d_5-31-19.csv")
```

Mean CT value by DNA isolation method. T tests and scatterplots comparing CT values.
```{r scatter}
######################################
# SWAB VS STOOL
######################################

#data frame for replicate 1 Ct values
ctDup1 <- d %>%
  select(personID, ctstool1, ctrectal1) %>%
  drop_na()
ctDup1$Duplicate <- rep(1,nrow(ctDup1))
names(ctDup1) <- c("ID", "Stool_CT", "Rectal_CT", "Duplicate")

#data frame for replicate 2 Ct values
ctDup2 <- d %>%
  select(personID, ctstool2, ctrectal2) %>%
  drop_na()
ctDup2$Duplicate <- rep(2,nrow(ctDup2))
names(ctDup2) <- c("ID", "Stool_CT", "Rectal_CT", "Duplicate")

#merge data frames
ctDupAll <- rbind(ctDup1, ctDup2)

#SCATTER PLOT
#############
scatterDup <- ggplot(ctDupAll, aes(Stool_CT, Rectal_CT)) + 
  geom_point(data = ctDupAll, color = "#543005",
             aes(shape = as.factor(Duplicate))) + 
  theme_light() + 
  labs(title = "", 
       x = "\nBulk stool CT value", 
       y = "\n\nRectal swab CT value\n") + 
  theme(axis.text = element_text(size = 10), 
        axis.title = element_text(size = 12), 
        plot.title = element_text(size = 12, 
                                  face = "bold", hjust = 0.5),
        legend.title = element_text(size = 12),
        legend.direction = "vertical",
        legend.position = c(0.84, 0.18),
        legend.box.background = element_rect(colour = "black")) +
  geom_smooth(method = lm, se = FALSE, color = "darkslategray", size = 0.5) +
  scale_y_continuous(limits=c(15, 35)) + 
  scale_x_continuous(limits=c(15, 35)) +
  scale_shape_manual(name = "Replicate", values = c(19, 17)) +
  geom_abline(intercept = 0, slope=1, linetype="dotted") + 
  annotate("text",
           x = 17,
           y = 32, 
           label = "paste(italic(R) ^ 2, \" = 0.171\")",
           size = 4,
           parse = TRUE) 
  #coord_fixed()

scatterDup
  
#ggsave("/Users/razel/Dropbox/Rectal swab STH/Analysis/Results/scatter_by_duplicate_7-22.jpg", 
        plot=last_plot(), width=6, height=6)

# R2 & P VALUE:
m1 <- lm(Rectal_CT ~ Stool_CT, data = ctDupAll)
summary(m1)
#Residual standard error: 3.719 on 49 degrees of freedom
#Multiple R-squared:  0.171,	Adjusted R-squared:  0.1541 
#F-statistic: 10.11 on 1 and 49 DF,  p-value: 0.002557

stargazer(ctDupAll, summary = TRUE, rownames = FALSE)
#shows ct means and sd per method

#p-value from t-test comparing overall means from all samples positive for swabs and stool
ctALL1 <- d %>%
  select(personID, ctstool1, ctrectal1) %>%
  drop_na()
  colnames(ctALL1) <- c("ID", "Stool_CT", "Rectal_CT")
ctALL2 <- d %>%
  select(personID, ctstool2, ctrectal2) %>%
  drop_na()
  colnames(ctALL2) <- c("ID", "Stool_CT", "Rectal_CT")
ctALL <- rbind(ctALL1, ctALL2)

ctALLTab <- ctALL %>%
  select(Stool_CT, Rectal_CT, ID) %>%
  gather(key="SampleType", value="CTvalue", -ID) %>%
  drop_na()

t.test(CTvalue ~SampleType, data = ctALLTab, paired=T)

#p-value from t-test comparing means by replicate group
t_CT1 <- d %>%
  select(ctstool1, ctrectal1, personID) %>%
  drop_na() %>%
  gather(key="SampleType", value="CTvalue", -personID) %>%
  mutate(SampleType = factor(SampleType)) %>%
  drop_na()

t.test(CTvalue ~SampleType, data = t_CT1, paired=T)

t_CT2 <- d %>%
  select(ctstool2, ctrectal2, personID) %>%
  drop_na() %>%
  gather(key="SampleType", value="CTvalue", -personID) %>%
  mutate(SampleType = factor(SampleType)) %>%
  drop_na()

t.test(CTvalue ~SampleType, data = t_CT2, paired=T)

group_by(t_CT1, SampleType) %>%
  summarise(
    count = n(),
    mean = mean(CTvalue, na.rm = TRUE),
    sd = sd(CTvalue, na.rm = TRUE)
  )


######################################
# STOOL 1 VS STOOL 2
######################################

#SCATTER PLOT
#############
scatterS <- ggplot(d, aes(ctstool1, ctstool2)) + 
  geom_point(color = "#35978f") + 
  theme_light() + 
  labs(title = "", 
       x = "\nBulk stool (1) CT value", 
       y = "\n\nBulk stool (2) CT value\n") + 
  theme(axis.text = element_text(size = 10), 
        axis.title = element_text(size = 12), 
        plot.title = element_text(size = 12, 
                                  face = "bold", hjust = 0.5), 
        legend.title = element_text(size = 12)) + 
  geom_smooth(method = lm, se = FALSE, color = "darkslategray", size = 0.5) +
  scale_y_continuous(limits=c(15, 35)) + 
  scale_x_continuous(limits=c(15, 35)) + 
  geom_abline(intercept = 0, slope=1, linetype="dotted") + 
  annotate("text",
           x = 17,
           y = 32, 
           label = "paste(italic(R) ^ 2, \" = 0.956\")",
           size = 4,
           parse = TRUE) 
    #coord_fixed()

#ggsave("/Users/razel/Dropbox/Rectal swab STH/Analysis/Results/scatter_stool_replicates.jpg", 
        plot=last_plot(), width=6, height=6)

# R2 & P VALUE:
m2 <- lm(ctstool2 ~ ctstool1, data = d)
summary(m2)
#Multiple R-squared:  0.9557,	Adjusted R-squared:  0.9542 
#F-statistic: 626.3 on 1 and 29 DF,  p-value: < 2.2e-16

d2 <- d %>%
  select(ctstool1, ctstool2, personID) %>%
  gather(key="SampleType", value="CTvalue", -personID) %>%
  drop_na()

group_by(d2, SampleType) %>%
  summarise(
    count = n(),
    mean = mean(CTvalue, na.rm = TRUE),
    sd = sd(CTvalue, na.rm = TRUE)
  )

t.test(CTvalue ~SampleType, data=d2, paired=T) # very high p-value indicates no difference 

######################################
# SWAB 1 VS SWAB 2
######################################

#SCATTER PLOT
#############
scatterR <- ggplot(d, aes(ctrectal1, ctrectal2)) + 
  geom_point(color = "#8c510a") + 
  theme_light() + 
  labs(title = "", 
       x = "\nRectal swab (1) CT value", 
       y = "\n\nRectal swab (2) CT value\n") + 
  theme(axis.text = element_text(size = 10), 
        axis.title = element_text(size = 12),
        plot.title = element_text(size = 12, 
                                  face = "bold", hjust = 0.5), 
        legend.title = element_text(size = 12)) + 
  geom_smooth(method = lm, se = FALSE, color = "darkslategray", size = 0.5) +
  scale_y_continuous(limits=c(15, 35)) + 
  scale_x_continuous(limits=c(15, 35)) + 
  geom_abline(intercept = 0, slope=1, linetype="dotted") + 
  annotate("text",
           x = 17,
           y = 32, 
           label = "paste(italic(R) ^ 2, \" = 0.665\")",
           size = 4,
           parse = TRUE) 
    #coord_fixed()

scatterR

#ggsave("/Users/razel/Dropbox/Rectal swab STH/Analysis/Results/scatter_rectal_replicates.jpg", 
        plot=last_plot(), width=6, height=6)

# R2 & P VALUE:
m3 <- lm(ctrectal2 ~ ctrectal1, data = d)
summary(m3)
#Multiple R-squared:  0.6649,	Adjusted R-squared:  0.6481 
#F-statistic: 39.68 on 1 and 20 DF,  p-value: 3.777e-06

d2 <- d %>%
  select(ctrectal1, ctrectal2, personID) %>%
  gather(key="SampleType", value="CTvalue", -personID) %>%
  drop_na()

group_by(d2, SampleType) %>%
  summarise(
    count = n(),
    mean = mean(CTvalue, na.rm = TRUE),
    sd = sd(CTvalue, na.rm = TRUE)
  )

t.test(CTvalue ~SampleType, data=d2, paired=T) # high p-value indicates no difference 

######################################
# COMBINED SCATTERPLOTS FOR PAPER
######################################

title3 <- ggdraw() + draw_label("", fontface = "bold")
scatterGrid3 <- plot_grid(scatterDup, scatterS, scatterR, labels = c("A", "B", "C"), nrow = 1, align = "h")
scatterGrid3 <- plot_grid(title3, scatterGrid3, ncol = 1, rel_heights = c(0.1, 1))
scatterGrid3
#save_plot("scatter_grid_horizontal.jpg", scatterGrid3, base_height = 6, base_width = 18)

title2 <- ggdraw() + draw_label("", fontface = "bold")
scatterGrid2 <- plot_grid(scatterS, scatterR, labels = c("A", "B"), nrow = 1, align = "h")
scatterGrid2 <- plot_grid(title2, scatterGrid2, ncol = 1, rel_heights = c(0.1, 1))
scatterGrid2
#save_plot("scatter_grid_horizontal2.jpg", scatterGrid2, base_height = 6, base_width = 12)
```


Boxplot comparing median and mean Ct values in stool samples where rectal was positive and where rectal was negative.
```{r box}
#USING ONLY REPLICATE ONE FOR BOXPLOT.

ggplot(d, aes(x=as.factor(rectal1_pos), y=ctstool1, fill = as.factor(rectal1_pos), na.rm=T)) +
  geom_boxplot() +
  theme_light() +
  stat_summary(fun.y=mean, geom="errorbar", 
               aes(ymax = ..y.., ymin = ..y.., color="mean"), 
               width = .75, 
               linetype = "dashed",
               color = "#01665e") +
  scale_color_discrete(name = "", labels = c("Mean")) + 
  labs (x = "\nPCR detection by rectal swab", 
        y="Mean stool sample Ct value\n", 
        title = "Figure 3. Mean and median Ct values in stool samples with \ncorresponding PCR positive and PCR negative rectal swabs\n") + 
  theme(axis.text = element_text(size = 8), 
        axis.title = element_text(size = 10), 
        plot.title = element_text(size = 10, 
                                  face = "bold", hjust = 0.5), 
        legend.title = element_text(size = 10,),
        legend.position = "none",
        panel.grid.major.x = element_blank()) +
  scale_fill_manual(values=c("#c7eae5", "#80cdc1"), 
                    name = "", 
                    labels = c("Negative", "Positive")) +
  scale_x_discrete(labels = c("0" = "Negative", "1" = "Positive"))

#ggsave("/Users/razel/Dropbox/Rectal swab STH/Analysis/Results/giardia_boxplot.jpg", 
        plot=last_plot(), width=6, height=6)

# T-TEST
group_by(d2, rectal1_pos) %>%
  summarise(
    count = n(),
    mean = mean(ctstool1, na.rm = TRUE),
    sd = sd(ctstool1, na.rm = TRUE)
  )

t.test(ctstool1 ~rectal1_pos, data=d)

#USING BOTH REPLICATE GROUPS FOR BOXPLOT.

ggplot(d, aes(x=as.factor(eitherRectalPos), y=ctstoolMean, fill = as.factor(eitherRectalPos), na.rm=T)) +
  geom_boxplot() +
  theme_light() +
  stat_summary(fun.y=mean, geom="errorbar", 
               aes(ymax = ..y.., ymin = ..y.., color="mean"), 
               width = .75, 
               linetype = "dashed",
               color = "#f5f5f5") +
  scale_color_discrete(name = "", labels = c("Mean")) + 
  labs (x = "\nPCR detection by rectal swab", 
        y="Mean stool sample Ct value\n", 
        title = "") + 
  theme(axis.text = element_text(size = 8), 
        axis.title = element_text(size = 10), 
        plot.title = element_text(size = 10, 
                                  face = "bold", hjust = 0.5), 
        legend.title = element_text(size = 10,),
        legend.position = "none",
        panel.grid.major.x = element_blank()) +
  scale_fill_manual(values=c("#80cdc1", "#35978f"), 
                    name = "", 
                    labels = c("Negative", "Positive")) +
  scale_x_discrete(labels = c("0" = "Negative", "1" = "Positive"))

#ggsave("/Users/razel/Dropbox/Rectal swab STH/Analysis/Results/giardia_boxplot_ALL.jpg", 
        plot=last_plot(), width=6, height=6)

# T-TEST
group_by(d2, rectal1_pos) %>%
  summarise(
    count = n(),
    mean = mean(ctstool1, na.rm = TRUE),
    sd = sd(ctstool1, na.rm = TRUE)
  )

t.test(ctstoolMean ~eitherRectalPos, data=d)
```

Restrict validity estimates to low CT values. Create CT quartile bar chart.
```{r quartile}
d$stoolct_quart <- ntile(d$ctstool1, 4)

ggplot(d, aes(x=ctstool1)) +
  geom_histogram() + 
  theme_minimal()

d2 <- d %>%
  #filter(ctstool1<30 | is.na(ctstool1))
  filter(stoolct_quart != 4 | is.na(ctstool1))

tab6 <- xtabs(~rectal2_pos + truePos, data=d2)
tab6
tab6 <- tab6[ 2:1, 2:1]
epi.tests(tab6)
epi.kappa(tab6)
#sensitivity with all swab positives vs true positives = 84% (0.71, 0.94)
#sensitivity with swab 1 positives vs true positives = 80% (0.65, 0.90)
#sensitivity with swab 2 positives vs true positives = 86% (0.67, 0.96)

# add ct values to model
d3 <- d %>%
  filter(truePos==1)

sens <- glm(rectal1_pos~1 + ctstool1, family = binomial(link="logit"), 
            data=d3)

coef(summary(sens))
#ctstool1 has a significant effect on whether rectal swab 1 was positive (p = 0.0001551446)

sens <- glm(eitherRectalPos~1 + ctstoolMean, family = binomial(link="logit"), 
            data=d3)

coef(summary(sens))
#ctstoolMean has a significant effect on whether either rectal swab was positive (p = 0.0002894428)

#generate quartiles

group_by(d, stoolct_quart) %>%
  summarize(meanCT = mean(ctstool1),
            min = min(ctstool1), 
            max = max(ctstool1))

d3 <- d %>%
  filter(stool1_ID != 484196 & stool1_ID != 644754 & stool1_ID != 861856) %>%
  filter(truePos ==1) %>%
  group_by(stoolct_quart) %>%
    summarize(nPos = sum(rectal1_pos, na.rm = TRUE),
            N = n(),
            N_notMiss = sum(!is.na(rectal1_pos)),
            PropPos = mean(rectal1_pos, na.rm = TRUE),
            sd = sd(rectal1_pos, na.rm=TRUE)) %>%
  mutate(SE = sd/sqrt(N_notMiss), lowerCI = PropPos - 1.96*SE, upperCI = PropPos + 1.96*SE)

#BAR CHART

ggplot(d3, aes(x=as.factor(stoolct_quart), y=PropPos, fill=as.factor(stoolct_quart))) +
  geom_bar(stat= "identity") +
  #scale_y_continuous(labels=scales::percent) +
  #geom_errorbar(aes(ymin=lowerCI, ymax=upperCI), width=.2,
                #position=position_dodge(.9)) +
  # scale_fill_brewer(palette="Paired", labels=c("Stool sample", "Rectal swab")) + 
  theme_minimal() +
  theme(panel.grid.major.x = element_blank()) +
  theme(axis.text = element_text(size = 8), 
        axis.title = element_text(size = 10), 
        plot.title = element_text(size = 10, 
                                  face = "bold", hjust = 0.5), 
        legend.title = element_text(size = 10)) +
  scale_y_continuous(labels = percent, limits = c(0,1), expand = c(0,0)) +
  labs (y= "Percent rectal swabs PCR positive\n", 
        x = "\nStool CT quartile", 
        title = "") +
  scale_fill_manual(name = "Stool CT value", 
                      labels = c("Q1: 13.6 - 21.4 ", 
                                 "Q2: 21.41 - 25.4 ", 
                                 "Q3: 25.41-30.2 ", 
                                 "Q4: 30.3 - 34.7"),
                      values = rev(BrBG4_turq_dark)) +
                      #Brown color pallete 
                      #values = rev(BrBG4_brown)) +
  geom_text(aes(label=c("14/14", "9/13", "10/14", "1/13")), 
            vjust=1.6, 
            color="white", 
            size=3.5)

#ggsave("/Users/razel/Dropbox/Rectal swab STH/Analysis/Results/ct_quartile.jpg", 
        plot=last_plot(), width=8, height=6)
```

Comparing CT values for rectal swabs and bulk stool by stool consistancy.
```{r ctConsist}
d4 <- d %>%
  select(ctstool1, ctrectal1, loosewat1, personID) %>%
  gather(key="SampleType", value="CTvalue", -personID, -loosewat1) %>%
  drop_na()

ggplot(d4, aes(x=loosewat1, y=CTvalue, fill=SampleType, na.rm=T)) +
geom_boxplot() +
  theme_light() +
  #stat_summary(fun.y=mean, geom="errorbar", 
               #aes(ymax = ..y.., ymin = ..y.., color="mean"), 
               #width = .75, 
               #linetype = "dashed",
               #color = "#f5f5f5") +
  scale_color_discrete(name = "", labels = c("Mean")) + 
  labs (x = "", 
        y="Cycle threshold value\n", 
        title = "Figure 5. Mean and median CT values by \nspecimen type and participant stool consistency\n") + 
  theme(axis.text = element_text(size = 8), 
        axis.title = element_text(size = 10), 
        plot.title = element_text(size = 10, 
                                  face = "bold", hjust = 0.5), 
        legend.title = element_text(size = 10,),
        legend.position = "bottom",
        panel.grid.major.x = element_blank()) +
  scale_fill_manual(values=c("#dfc27d", "#35978f"), 
                    name = "", 
                    labels = c("Rectal swab", "Bulk stool")) +
  scale_x_discrete(labels = c("loose/watery" = "Loose or watery stool", "formed" = "Formed stool"))

#ggsave("/Users/razel/Dropbox/Rectal swab STH/Analysis/Results/consistency_boxplot.jpg", 
        plot=last_plot(), width=6, height=6)
```

Age and CT value plots.
```{r agePlots}
ggplot(d, aes(x=ageY, y=ctstool1)) +  
  geom_point() +
  geom_smooth(method = lm, se = FALSE, color = "darkslategray", size = 0.5) +
  theme_minimal()

d5 <- d %>%
  select(stool1_pos, rectal1_pos, ageY, personID) %>%
  gather(key="SampleType", value="Pos", -personID, -ageY) 

ggplot(d5, aes(x=ageY, y=Pos, color=SampleType)) +  
  geom_smooth() +
  theme_minimal()

ggplot(dTest, aes(x=ageY, y=ctstoolMax)) +
  geom_smooth() +
  theme_minimal()

d6 <- d %>%
  select(ctstool1, ctrectal1, ageY, personID) %>%
  gather(key="SampleType", value="CTvalue", -personID, -ageY) 
 
ggplot(d6, aes(x=ageY, y=CTvalue, color=SampleType)) +  
  geom_smooth(fill="#D0DEDC") +
  #geom_point() +
  theme_light() +
  theme(axis.text = element_text(size = 8), 
        axis.title = element_text(size = 10), 
        plot.title = element_text(size = 10, 
                                  face = "bold", hjust = 0.5), 
        legend.title = element_text(size = 10,),
        legend.position = "right",
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()) +
  labs (x = "\nAge in years", 
        y="Cycle threshold value\n") + 
        #title = "Real-time PCR CT value by specimen type and age of child\n") + 
  scale_fill_manual(values=c("#dfc27d"), 
                    name = "", 
                    labels = c("Rectal swab", "Bulk stool")) + 
  scale_color_manual(name = "Specimen type", 
                     labels = c("Rectal swab", "Bulk stool"), 
                     values = c("#8c510a", "#35978f"))

#ggsave("/Users/razel/Dropbox/Rectal swab STH/Analysis/Results/ct_smooth.jpg", 
        plot=last_plot(), width=8, height=6)
```
